### 2022-03-24 Writing/Palaeontology/draft.tex
Updated discussion

\section{Discussion}

\subsection{Effect of character differences on topology}
We find that there is a significant influence of morphological character correlation on inferring tree topology.
When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
Throughout our simulation, the ``minimised'' character difference scenario (i.e. increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario led to slightly worse trees (Fig. \ref{Fig:RF_results_best}).
These results suggest that, in matrices with high character differences, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.

However, this effect of increasing or reducing character difference did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
Furthermore, these results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
Furthermore, fewer characters were replaced, on average, in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).
This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.

\subsubsection{Number of characters and taxa}
We found an effect of the number of characters on the ability to recover the ``best'' topology with different levels of character correlations only between the two extreme number of characters (100 and 1000; Tables \ref{Tab_pooledscharacters_test}).
This is expected since a higher number of characters can result in better topologies with a higher probability of ``noisy'' signal to cancel (\textit{c.f.} in smaller matrices).
While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}).
This can be explained partly by our method of comparing trees using the Normalised Tree Similarity \citep[$NTS$;][]{Bogdanowicz2012} which corrects for the fact that topological differences are proportional to the number of taxa considered.

\subsubsection{Effect of character differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of each method to recover the ``correct'' topology has been recently discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (using some M$k$-based characters) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[using matrices with a $CI>0.26$;][]{OReilly20160081} could also favour maximum parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

\subsection{Distinction between different character correlations}
Here we mention three different types of character correlations but evolutionary biologists are mainly interested in intra-organismal and evolutionary correlations (e.g. in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters (to avoid creating circular and biased reasoning).
This creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.

@@

\subsection{Potential applications}
Effectively, our simulation protocol bootstraps our data ``with bias''.
In the ``randomised'' scenarios the data is simply randomly bootstrapped.
However, in the ``minimised'' and ``maximised'' scenario, we remove the characters with the lowest/highest overall character difference.
For example, in the ``maximised'' scenario, we randomly remove some characters that are strongly correlated with each other and randomly resample from the characters left.

In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when maximising or randomising the character differences.
Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling
will be sufficient to obtain a robust topology.
In matrices with more taxa, however, the ``minimised'' scenario resulted in better topological recovery.
Applying this type of bootstraps that minimised character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters. 
Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the data.
In fact, regions of the matrix where characters have a consistent low difference can reflect that they are coding for the same information.

\subsection{Limitations}
First, simulating evolutionary history is complex.
Not only because the models we're using to infer phylogenies are ever-improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
However, we do not compare the ``maximised'', ``minimised'' and ``randomised'' to the ``true'' tree but rather to the ``normal'' tree.
This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.

Second, comparing phylogenetic inference methods is not trivial.
As mentioned above, both maximum parsimony and Bayesian inference have similar outputs nut vastly differ in how optimality is measured.
There are also difficulties in summarising both methods with consensus trees \cite{oReilly2017efficacy}.
However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.

@@

We found that character correlation has a strong effect on recovering the correct topology: matrices with high character differences always performed more poorly at recovering the correct topology than trees with low or randomised character differences.
This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
However, in empirical datasets, if character difference is not driven by selection (e.g. pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.
We recommend that all morphological phylogenetic analyses should report the level of character difference so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.

>>>>

%---------------------------------------------
%
%       DISCUSSION
%
%---------------------------------------------
\section{Discussion}

\subsection{Effect of character differences on topology}
We find that there is a significant influence of morphological character correlation on inferring tree topology.
When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
Throughout our simulation, the maximised character correlation scenario (i.e. decreased character difference) tended to result in trees slightly closer to the tree inferred from the unperturbed matrix and the minimised character correlation led to slightly worse trees (Fig. \ref{Fig:RF_results_best}).
These results suggest that, in matrices with low character correlations, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.

However, this effect of increasing or reducing character correlation did not appear to affect taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped to much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
Furthermore, these results needs to be contrasted with the fact that the trees inferred from the maximised character correlation matrices do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
Furthermore, fewer characters were replaced, on average, in the maximised correlation matrices compared to both the minimised correlation and the randomised matrices (supplementary material 3, Figs 7, 8 and 9) although this did not change the way the average character differences varied across the matrices (supplementary material 3, Figs 4, 5 and 6).

\subsubsection{Number of characters and taxa}
We found an effect of the number of characters on the ability to recover the topology from the unperturbed matrices with different levels of character correlations only between the two extreme number of characters (100 and 1000; Tables \ref{Tab_pooledscharacters_test}).
This is expected since a higher number of characters can result in better topologies with a higher probability of ``noisy'' signal to cancel (\textit{c.f.} in smaller matrices).
While matrices with more characters or taxa tended to be better resolved than matrices with fewer characters or taxa, the effect of character correlation on topology is independent of the size of the matrix (Tables \ref{Tab_pooledscharacters_test} and \ref{Tab_pooledstaxa_test}).
This can be explained partly by our method of comparing trees using the Normalised Tree Similarity \citep[$NTS$;][]{Bogdanowicz2012} which corrects for the fact that topological differences are proportional to the number of taxa considered.

\subsubsection{Effect of character differences on the inference method}
When considering the pooled effect of the tree inference method, we only detected a significant difference between the Bayesian and the maximum parsimony trees in terms of clade conservation but none in terms of taxa placement (Table \ref{Tab_pooledsmethods_test}).
The difference in the ability of each method to recover the topology inferred from the unperturbed matrices has been discussed with some indications that Bayesian inference will outperform parsimony when analysing discrete morphological characters alone (\citealt{wrightbayesian2014,OReilly20160081,puttick2017uncertain}; but see \citealt{spencerefficacy2013,goloboff2017weighted}).
In this study, it is possible that our simulation protocol for generating the characters (using some M$k$-based characters) could slightly favour Bayesian inference over maximum parsimony, however, our protocol for selecting matrices \citep[using matrices with a $CI>0.26$;][]{OReilly20160081} could also favour maximum parsimony analysis.
It was however not the purpose of this study to compare the overall performance of both methods but rather to measure the effect of character correlation on each of those methods separately.

\subsection{Distinction between different character correlations}
Here we mention three different types of character correlations but \textcolor{blue}{only look at raw (logical) correlation and do not distinguish between the different character dependencies mentioned in the introduction}.
However, it is likely that evolutionary biologists are mainly interested in intra-organismal and evolutionary correlations (e.g. in evo-devo \citealt{goswami2006morphological}; or in macroevolution \citealt{fitzjohn2014much}).
These two types of correlations can only be studied \textit{a posteriori} with a phylogenetic hypothesis and should not be used \textit{a priori} as a criterion to select characters (to avoid creating circular and biased reasoning).
This creates a trade-off between: (1) coding fewer characters (stochastically reducing \textit{a priori} correlation) but making the \textit{a posteriori} correlation more dependent on the coding;
and (2) coding more characters (increasing \textit{a priori} dependence) but allowing the \textit{a posteriori} correlation to be less dependent on the coding correlations.

@@

\subsection{Potential applications}
Effectively, our simulation protocol bootstraps data ``with bias''.
In the randomised matrices the data is simply randomly bootstrapped.
However, in the matrices with maximised and minimised correlation, we remove the characters with the highest/lowest overall character difference.
For example, in the matrices with minimised correlation, we randomly remove some characters that are strongly correlated with each other and randomly resample from the characters left.

In rather small matrices (25 $\times$ 100), there was no significant difference in terms of recovering the right topology when minimising or randomising the character correlation.
Since many discrete morphological matrices are of similar size \citep{guillerme2016assessment} a simple bootstrap re-sampling will be sufficient to obtain a robust topology.
In matrices with more taxa, however, maximising the character correlation resulted in better topological recovery.
Applying this type of bootstraps that decreased character difference by biasing the random sampling could thus help in understanding which characters generate more homoplasy, which nodes in the tree are conflictual or which clades might be most dependent on highly correlated characters. 
Overall, we suggest measuring (and reporting) the levels of character difference in a matrix to help understand potential caveats in the data.
In fact, regions of the matrix where characters have a consistent low difference can reflect that they are coding for the same information.

\subsection{Limitations}
First, simulating evolutionary history is complex.
Not only because the models we're using to infer phylogenies are ever-improving \citep[e.g.][]{heath2014fossilized,Wright01072016} but also because generalising morphological evolution across vastly different organisms is probably impossible \citep[see constrasted discussions from][]{GoloboffEmpirical,OReillyEmpirical}.
However, we do not compare the trees resulting from the \textcolor{blue}{maximised, minimised or randomised correlation matrices to the starting tree but rather to the tree inferred from the unperturbed matrix.}
This allows us to reduce the caveats from our simulations on the effect of character correlation since we only compare the simulation end products to the other outputs rather than to the simulation inputs.

Second, comparing phylogenetic inference methods is not trivial.
As mentioned above, both maximum parsimony and Bayesian inference have similar outputs but vastly differ in how optimality is measured.
There are also difficulties in summarising both methods with consensus trees \citep{oReilly2017efficacy}.
However, we want to point out again that here we are not comparing the methods to each other \textit{per se} but rather how they both, individually, react to an increase or decrease of correlated characters.

\textcolor{blue}{Third, by not focusing on modeling the sources of character correlation specifically (see section ``Distinction between different character correlations'' above), we do not allow to distinguish the specific effect of each correlation for downstream analyses (see \citealt{sansom2017dental}f or such an application on dental morphological data).
However, these sources are often not available to authors, either because of the data availability or the resources available to the workers (i.e. time and money) and thus we here propose an approach showing the effect of character correlation when the sources of correlation are effectively unknown.}

@@

We found that character correlation has an effect on recovering the correct topology: matrices with low character correlation always performed more poorly at recovering the correct topology than matrices with high or randomised character correlation.
This is especially true in large matrices where we advise to reduce character difference by using bootstrapping with ``bias'' whereas this may not be major problem in modest size matrices (25 taxa; 100 to 350 characters).
However, in empirical datasets, if character difference is not driven by selection (e.g. pleiotropy) correlation is likely be cancelled out if the correlated characters are randomly distributed with respect to traits.
We recommend that all morphological phylogenetic analyses should report the level of character correlation or differences so that we can begin to detect and address the problem of coding dependence on inflating our confidence in our phylogenetic reconstructions.
